### 2022-09-20 analysis/supplementary-materials/PetrologicalAnalysis.Rmd

---
title: "Thin section analysis"
author: "Joana Belmiro"
date: "5/20/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

```{r library, include=FALSE}
library(knitr)
library(dplyr)
library(readr)
library(kableExtra)
library(tidyr)
```

```{r readbd, include=FALSE}

TSdb <- read_csv(here::here('analysis/data/raw_data/PetrographyDB.csv')) # RM types file

```

# Geological samples

## SP6

Sample `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP6") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP6") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP6") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ORTHOCHEM2TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP6") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP6") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP6") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP6") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP6") %>% select(BIOCLAST2))`. There is `r as.character(TSdb %>% filter(ID == "SP6") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP6") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP6") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP7

Sample `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP7") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP7") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP7") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ORTHOCHEM3TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP7") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP7") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP7") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP7") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP7") %>% select(BIOCLAST2))`. There is `r as.character(TSdb %>% filter(ID == "SP7") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP7") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP7") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP9

Sample `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP9") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP9") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP9") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ORTHOCHEM2TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP9") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP9") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP9") %>% select(BIOCLAST1))`. There is `r as.character(TSdb %>% filter(ID == "SP9") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP9") %>% select(POROSITYTYPE))` type. Other sedimentary features are a `r as.character(TSdb %>% filter(ID == "SP9") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP10

Sample `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP10") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP10") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP10") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ORTHOCHEM2TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ALLOCHEM1))`; `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ALLOCHEM2FREQ))` `r as.character(TSdb %>% filter(ID == "SP10") %>% select(ALLOCHEM2))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP10") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP10") %>% select(BIOCLAST1))`. There is `r as.character(TSdb %>% filter(ID == "SP10") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP10") %>% select(POROSITYTYPE))` type. Other sedimentary features are a `r as.character(TSdb %>% filter(ID == "SP10") %>% select(SEDIMENTARYSTRUCTURES))`. `r as.character(TSdb %>% filter(ID == "SP10") %>% select(OBSERVATIONS))`.

## SP14

Sample `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP14") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP14") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP14") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ORTHOCHEM3TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP14") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP14") %>% select(BIOCLAST3))`. `r as.character(TSdb %>% filter(ID == "SP14") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP14") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP14") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP14") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP18

Sample `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP18") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP18") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP18") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM3TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM4FREQ))` of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM4))`, of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ORTHOCHEM4TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ALLOCHEM1))`; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ALLOCHEM2FREQ))` `r as.character(TSdb %>% filter(ID == "SP18") %>% select(ALLOCHEM2))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP18") %>% select(BIOCLAST3))`. There is `r as.character(TSdb %>% filter(ID == "SP18") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP18") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP18") %>% select(SEDIMENTARYSTRUCTURES))`. `r as.character(TSdb %>% filter(ID == "SP18") %>% select(OBSERVATIONS))`.

## SP21

Sample `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP21") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP21") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP21") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM3TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM4FREQ))` of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM4))`, of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ORTHOCHEM4TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP21") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP21") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP21") %>% select(BIOCLAST1))`. `r as.character(TSdb %>% filter(ID == "SP21") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP21") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP21") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP21") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP23

Sample `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP23") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP23") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP23") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM3TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM4FREQ))` of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM4))`, of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ORTHOCHEM4TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST3))`; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST4FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST4))`; `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST5FREQ))` `r as.character(TSdb %>% filter(ID == "SP23") %>% select(BIOCLAST5))`. `r as.character(TSdb %>% filter(ID == "SP23") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP23") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP23") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP23") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP27

Sample `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP27") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP27") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP27") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ORTHOCHEM3TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP27") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP27") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP27") %>% select(BIOCLAST1))`.  `r as.character(TSdb %>% filter(ID == "SP27") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP27") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP27") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP27") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP32

Sample `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP32") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP32") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP32") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ORTHOCHEM2TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ALLOCHEM1))`; `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ALLOCHEM2FREQ))` `r as.character(TSdb %>% filter(ID == "SP32") %>% select(ALLOCHEM2))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP32") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP32") %>% select(BIOCLAST1))`. `r as.character(TSdb %>% filter(ID == "SP32") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP32") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP32") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP32") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP34

Sample `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP34") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP34") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP34") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM3TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM4FREQ))` of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM4))`, of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ORTHOCHEM4TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST3))`; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST4FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST4))`; `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST5FREQ))` `r as.character(TSdb %>% filter(ID == "SP34") %>% select(BIOCLAST5))`. `r as.character(TSdb %>% filter(ID == "SP34") %>% select(OBSERVATIONS))`. There is `r as.character(TSdb %>% filter(ID == "SP34") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP34") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP34") %>% select(SEDIMENTARYSTRUCTURES))`.

## SP39

Sample `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP39") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP39") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP39") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ORTHOCHEM3TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ALLOCHEM1))`; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ALLOCHEM2FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(ALLOCHEM2))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST3))`; `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST4FREQ))` `r as.character(TSdb %>% filter(ID == "SP39") %>% select(BIOCLAST4))`. There is `r as.character(TSdb %>% filter(ID == "SP39") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP39") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP39") %>% select(SEDIMENTARYSTRUCTURES))`. `r as.character(TSdb %>% filter(ID == "SP39") %>% select(OBSERVATIONS))`.

## SP40

Sample `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP40") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP40") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP40") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM2TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM3FREQ))` of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM3))`, of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ORTHOCHEM3TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ALLOCHEM1))`; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ALLOCHEM2FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(ALLOCHEM2))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST2))`; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST3FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST3))`; `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST4FREQ))` `r as.character(TSdb %>% filter(ID == "SP40") %>% select(BIOCLAST4))`. There is `r as.character(TSdb %>% filter(ID == "SP40") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP40") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP40") %>% select(SEDIMENTARYSTRUCTURES))`. `r as.character(TSdb %>% filter(ID == "SP40") %>% select(OBSERVATIONS))`.

## SP42

Sample `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ID))` is a `r as.character(TSdb %>% filter(ID == "SP42") %>% select(LITHOLOGY))`, characterized by a `r as.character(TSdb %>% filter(ID == "SP42") %>% select(TEXTURE))` texture and a `r as.character(TSdb %>% filter(ID == "SP42") %>% select(MICROSTRUCTURE))` microstructure. `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEMCOUNT))` orthochems were identified: `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM1FREQ))` of `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM1))`, of `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM1TYPE))` type; `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM2FREQ))` of `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM2))`, of `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ORTHOCHEM2TYPE))` type. Allochems are `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ALLOCHEM1FREQ))` `r as.character(TSdb %>% filter(ID == "SP42") %>% select(ALLOCHEM1))`. Bioclasts are `r as.character(TSdb %>% filter(ID == "SP42") %>% select(BIOCLAST1FREQ))` `r as.character(TSdb %>% filter(ID == "SP42") %>% select(BIOCLAST1))`; `r as.character(TSdb %>% filter(ID == "SP42") %>% select(BIOCLAST2FREQ))` `r as.character(TSdb %>% filter(ID == "SP42") %>% select(BIOCLAST2))`. There is `r as.character(TSdb %>% filter(ID == "SP42") %>% select(TOTALPOROSITY))` of total porosity of `r as.character(TSdb %>% filter(ID == "SP42") %>% select(POROSITYTYPE))` type. Other sedimentary features are `r as.character(TSdb %>% filter(ID == "SP42") %>% select(SEDIMENTARYSTRUCTURES))`.