### 2022-10-04 analysis/supplementary-materials/SP6.Rmd
- Uploading the word draft with corrections

---
title: "Untitled"
author: "Joana Belmiro"
date: '2022-10-01'
output: pdf_document
---

# Sample ID and provenance

Sample ID: SP6_CSV

Lithology: Chert

Collection: LusoLit

Outcrop: Cabo S. Vicente

Unit/facies: Lower Jurassic

***

## Macroscopic description

### **Color**

The color distribution is Single and the color is Light gray (2.5Y 7/2).

### **Fabric**

The luster is Dull and the translucency is Opaque. The feel is Semi-smooth to Rough and the grain is fine. The distribution of the fabric is Even.

### **Inclusions and fossil content**

### **Cortex**

The cortex is Outcrop type with a Gradual transition.

### **Quality**

The fracture is Conchoidal with Fractures. The knapping quality is Medium.

### **Observations**

***

## Petrography analysis form

### Textural classification

**Texture**
Mudstone

**Microstructure**
Massive

***

### Composition

| Orthochem       | Type | %   | Description                             |
| --------------- | ---- | --- | --------------------------------------- |
| MiC quartz (gr) | SE   | 90  |                                         |
| Dolomite        | SE   | 10  | Dolomite crystals have different sizes. |

| Allochem | Freq     | Description |
| -------- | -------- | ----------- |
| Oxides   | Uncommon |             |

| Bioclasts  | Freq     | Description |
| ---------- | -------- | ----------- |
| Ghosts     | Uncommon |             |
| Echinoderm | Uncommon |             |

### Other textural characteristics

**Porosity**

Total porosity (%): 5

Porosity: Vuggy

***

### Photos

| Image ID | Aug. | Description                                                           |
| -------- | ---- | --------------------------------------------------------------------- |
| SP6_001  | 10x  | General view of oxides and a possible Echinoderm spine.               |
| SP6_002  | 10x  | Ghosts with quartz filling.                                           |
| SP6_003  | 10x  | Limit between chert and dolomitic limestone (cortex.)                 |
| SP6_004  | 10x  | Possible Echinoderm spine.                                            |
| SP6_005  | 20x  | Detail of rhomboid dolomite crystals close to the edge of the sample. |

***
### Observations

- Ghosts are filled with microcrystalline quartz and poorly preserved.
- Dolomite is visible in the chert area closer to the cortex area.

***
### Analysis information

- Analyst: JB

- Date: 02.23.2022

- Equipment: Leica DM2500 P