### 2022-10-08 analysis/supplementary-materials/thinsection_list.Rmd
Supplementary data: - geology references - macroscopic analysis table - thin section list

---
title: "List of thin sections"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

| Sample ID | Type               | Laboratory | Outcrop          | Sample collection |
| --------- | ------------------ | ---------- | ---------------- | ----------------- |
| SP6       | Covered            | UB         | Cabo S. Vicente  | 2021              |
| SP7       | Covered            | UB         | Cabo S. Vicente  | 2021              |
| SP9       | Covered            | UB         | Ponta dos Altos  | 2021              |
| SP10      | Covered            | UB         | Ponta dos Altos  | 2021              |
| SP14      | Covered            | UB         | Praia Belixe     | 2021              |
| SP15      | Covered            | UB         | Praia Belixe     | 2021              |
| SP18      | Covered            | UB         | Praia Belixe     | 2021              |
| SP21      | Covered            | UB         | Belixe Sul       | 2021              |
| SP23      | Covered            | UB         | Belixe Norte     | 2021              |
| SP24      | Covered            | UB         | Cabo S. Vicente  | 2021              |
| SP27      | Covered            | UB         | Cabo S. Vicente  | 2021              |
| SP28      | Covered            | UB         | Aspa             | 2021              |
| SP32      | Covered            | UB         | Foz dos Fornos   | 2021              |
| SP33      | Covered            | UB         | Foz dos Fornos   | 2021              |
| SP34      | Covered            | UB         | Ponta dos Altos  | 2021              |
| SP34      | Covered            | UB         | Ponta dos Altos  | 2021              |
| SP36      | Covered            | UB         | Ponta da Atalaia | 2021              |
| SP39      | Covered            | UB         | Andorinha        | 2021              |
| SP40      | Covered            | UB         | Ferrel           | 2021              |
| SP42      | Covered            | UB         | Boca do Rio      | 2021              |
| SP59      | Covered            | TSL        | Jordana          | 2022              |
| SP61      | Covered            | TSL        | Peral            | 2022              |
| SP56      | Covered            | TSL        | Jordana          | 2022              |
| SP54      | Covered            | TSL        | Guilhim          | 2022              |
| SP53      | Covered            | TSL        | Guilhim          | 2022              |
| SP58      | Covered            | TSL        | Jordana          | 2022              |
| SP55      | Covered            | TSL        | Caliços          | 2022              |
| SP50      | Covered            | TSL        | Casal da Colina  | 2022              |
| SP52      | Covered            | TSL        | Casal da Colina  | 2022              |
| RT82      | Covered            | TSL        | Praia Belixe     | <2021             |